---
title: "Output data for simulation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Output data for simulation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Table of Contents

[User Guide content](A_User_Guide.html)

[Theory of the model](B_Theory.html)

[Quick start guide](C_Quick_start.html)

[Simulation preparation](D_Simulation.html)

[Input data](E_Input.html)

[Output data](F_Output.html)

[Post simulation analisys](G_Analisys.html)

[VAF calculation](H_VAF.html)

[Figures](I_Figures.html)

[Relation to experimental measurements](J_Relation.html)

[Versions](K_Versions.html)

[References](L_References.html)


---


The output data consists of several files after the simulation. 

## "log.txt" file
The file **"log.txt"** contains information about probabilities and file names. These variables are explained in the ["Inputs"](E_Input.html). 

```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Output/log.txt",header = FALSE, sep = "\t", col.names = c("Variable","Value"), stringsAsFactors = FALSE )
x[is.na(x)] <- ""
x[ c(6,7) , 1 ] =  c( 'E0' , 'F0' )
knitr::kable(x, align = "c", caption = "**The log.txt file.** Example of log.txt file.")
```


## "geneout.txt" file
The file **"geneout.txt"** contains input information about the weights that connect the hallmarks and genes, which are defined by the user. These variables also are explained in the ["Inputs"](E_Input.html). 

```{r, echo=FALSE, results='asis'}
x <- read.csv(file = '../inst/extdata/Output/geneout.txt', header = TRUE, sep = "\t", col.names = c("Gene_name","Hallmark_name", "Weight", "Weight_CF", "Suppressor_or_oncogene"), stringsAsFactors = FALSE )
x[is.na(x)] <- ""
knitr::kable(x, align = "c", caption = "**The geneout.txt file.** Given below is an example of the geneout.txt file.")
```

'Weight_CF' column has weights of hallmark multiplied by CF or compaction factor for each hallmark respectively. 


## "cloneout.txt" file

The file **"cloneout.txt"** contains the results of the simulation and includes the evolution data: all the output data for each clone at each time step (only the first 10 lines are presented): 

```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Output/cloneout.txt",header = TRUE, sep = "\t", nrows = 10, stringsAsFactors = FALSE )
x <- x[,c(1:13)] 
# library(stringr)
# library(tidyr)
x[is.na(x)] <- ""
library(stringr)
x[] <- lapply(x, str_trunc, 6, ellipsis = "")
names( x )[ 6 ] = 'Birth_time'
knitr::kable(x, align = "c", caption = "**The Output data from cloneout.txt file.** Example of output data for all clones. Columns are from 1 to 13.")
```

1. **Time** - the time step, e.g., 1, 50.
2. **N\_cells** - the number of cells in this clone, e.g. 1000, 2.
3. **AvgOrIndx** - "avg" or "index": "avg" is for a line with averaged values across different (index) lines at the same time step; "index" shows the cell's index at the current time step,  e.g., avg, 4,7.
4. **ID** - the unique ID of clone, e.g., 1, 50.
5. **Parent_ID** - the parent index,  e.g., 0, 45.
6. **Birth_time** - the time step of the clone's birth,  e.g., 0, 5.
7. **c** - the counter of cell divisions for the clone, it equals average counter across all the cells in the clone.
8. **d**  - the probability of division for the cell, e.g., 0.1, 0.8 [per time-step].
9. **i**  - the probability of immortalization for the cell, e.g., 0.1, 0.8 [per time-step].
10. **im**   - the probability of invasion/metastasis for the cell, e.g., 0.1, 0.8 [per time-step].
11. **a**   - the probability of apoptosis for the cell, e.g., 0.1, 0.8 [per time-step].
12. **k**   - the probability of death due to the environment, e.g., 0.1, 0.8 [per time-step].
13. **E**  - the E coefficient for the function of the division probability, e.g., 10^4, 10^5.

 

```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Output/cloneout.txt",header = TRUE, sep = "\t", nrows = 18, stringsAsFactors = FALSE )
x[is.na(x)] <- ""
x <- x[,c(1,3,14:24)] 
library(stringr)
x[,c(1:11,13)] <- lapply(x[,c(1:11,13)], str_trunc, 7, ellipsis = "")
#names(x)[10] = 'driver_genes'
knitr::kable(x, align = "c", caption = "**cloneout.txt**. Columns are from 14 to 24.")
```

14. **N_normal** - the number of normal cells at this time step, e.g., 134, 5432.
15. **Nmax** - the theoretically maximal number of primary tumor cells, e.g., 10000, 5000.
16. **N_primary** - the number of primary tumor cells at this time step, e.g., 134, 5432.
17. **N_metastatic** - the number of metastatic cells at this time step, e.g., 16, 15439.
18. **Ha** - the value of the hallmark "Apoptosis" for the cell, e.g., 0.1, 0.4444.
19. **Him** - the value of the hallmark "Invasion / Metastasis" for the cell, e.g., 0.1, 0.4444.
20. **Hi** - the value of the hallmark "Immortalization" for the cell, e.g., 0.1, 0.4444.
21. **Hd** - the value of the hallmark "Growth / Anti-growth" for the cell, e.g., 0.1, 0.4444 .
22. **Hb** - the value of the hallmark "Angiogenesis" for the cell, e.g., 0.1, 0.4444 .
23. **type** - the type of the cell: 'normal' or 'primary' or 'metastatic'.
24. **mut_den** - the density of mutations for the cell, it equals to ratio a number of mutated driver genes to a number of all the genes, e.g., 0, 0.32.

 
```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Output/cloneout.txt",header = TRUE, sep = "\t", nrows = 18, stringsAsFactors = FALSE)
x[is.na(x)] <- ""
x <- x[,c(1,3,25:28)] # ncol(x)] 
knitr::kable(x, align = "c", caption = "**cloneout.txt**. Columns are from 25 to 28.")
```
25. **driver_genes** - the binary numbers indicate the driver mutation at the gene related to order of genes in onco as well as order of the next columns with genes' names, e.g., '1 0 0 0' means that the first gene has a driver mutation and other genes have no.
26. **passenger_genes** - the binary numbers indicate the passenger mutation at the gene related to order of genes in onco as well as order of the next columns with genes' names, e.g., '0 0 1 0' means that the third gene has a passanger mutation and other genes have no.
27. **PointMut_ID** - the index of data row for point mutation data frame saved at the end of simulation in the file **Point_mutations.txt**, e.g., 23, 32.
28. **CNA_ID** - the index of data row for CNA data frame saved at the end of simulation in the file **CNA.txt**, e.g., 44, 21.

There are two columns (25th and 26th) with the indexes of point mutations and CNAs. 
Each index corresponds to index in the related data frames for point mutations and for CNAs represented in additional tables below.

 
```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Output/cloneout.txt",header = TRUE, sep = "\t", nrows = 18, stringsAsFactors = FALSE)
x[is.na(x)] <- ""
x <- x[,c(1,3,29:32)] # ncol(x)] 
knitr::kable(x, align = "c", caption = "**cloneout.txt**. Columns are from 29 to 32.")
```

29-32. **Chr1_CDS_(gene's name)**, for example **Chr1_CDS_APC** - the length of CDS for each gene in the order of names of genes for FIRST parental chromosome of a clone. The CDS length of genes for second parental chromosome can be different in principle. The point mutation is proportional to **CDS_(gene's name)** od each chromosome. 


 
```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Output/cloneout.txt",header = TRUE, sep = "\t", nrows = 18, stringsAsFactors = FALSE )
x[is.na(x)] <- ''
x <- x[, c( 1,3, 33:36 ) ] 
library(stringr)
x[] <- lapply(x, str_trunc, 8, ellipsis = "")
knitr::kable(x, align = "c", caption = "**cloneout.txt**. Columns are from 33 to 36.")
```

33-36. **Chr1_Len_(gene's name)**, for example **Chr1_Len_APC** - the length of gene in the order of names of genes for ONLY FIRST chromosome of a clone. The length of genes for second chromosome can be different in principle. The CNA mutation is proportional to **Len_(gene's name)**. 


 
```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Output/cloneout.txt",header = TRUE, sep = "\t", nrows = 18, stringsAsFactors = FALSE )
x[is.na(x)] <- ''
x <- x[, c( 1,3, 37:40 ) ] 
library(stringr)
x[] <- lapply(x, str_trunc, 8, ellipsis = "")
knitr::kable(x, align = "c", caption = "**cloneout.txt**. Columns are from 37 to 40.")
```

37. **Chr1_p0** - the probability that during a trial, a cell of the clone has **NO** mutation [per time-step]. Applied to all cells in the clone.

38. **Chr1_prob_point_mut** - the **conditional** probability that if cell will have a mutation it should be a **point mutation**. 

39. **Chr1_prob_del** - the **conditional** probability that if cell will have a mutation it should be a **deletion**. 

40. **Chr1_prob_dup** - the **conditional** probability that if cell will have a mutation it should be a **duplication**. 


Last columns (from 41 to the end) show the data related to the **SECOND** parental chromosome with the same set of data as for the first parental chromosome.

 
```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Output/point_mutations.txt",header = TRUE, sep = "\t", stringsAsFactors = FALSE)
# x[is.na(x)] <- ""
x = x[ 1 : 20, ]
knitr::kable(x, align = "c", caption = "Point mutation data frame which will be saved to the file **point_mutations.txt** at the end of simulation.")
```


1. **PointMut_ID** - ID of point mutation, first ID is related to variant allele 'B' and same *second* ID - to the original allele A. 
2. **Parental_1or2** - indicates either of the two parental chromosomes. 
3. **Chr** - name of a chromosome.
4. **Ref_pos** - the reference position of an allele. The reference position is on the coordinate system of the human reference genome. 
5. **Phys_pos** - the physical position of an allele. The physical length of a (parental) chromosome is extended or shrunk by CNA duplications or deletions, respectively. When a duplication happens, the reference position is divided into two or more physical positions, which are represented by multiple elements in a vector. When a deletion happens and the allele is lost, the lost is represented by “-” on the coordinate system of physical positions. 
6. **Delta** - difference between the reference and physical positions.
7. **Copy_number** - the copy number of an allele. 
8. **Gene_name** - the name of a gene.
9. **MalfunctionedByPointMut** - logical indicator of whether or not the gene is malfunctioned by the point mutation. 
10. **mut_order** - indicator of mutation order in the simulation, it's used to detect order of mutations in the clone at each chromosome.



```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Output/CNA_mutations.txt",header = TRUE, sep = "\t", stringsAsFactors = FALSE)
# x[is.na(x)] <- ""
x = x[ 1 : 20, ]
knitr::kable(x, align = "c", caption = "**CNA mutation data frame which will be saved to the file **CNA_mutations.txt** at the end of simulation.")
```

1. **CNA_ID** - ID of CNA. 
2. **Parental_1or2** - indicates either of the two parental chromosomes. 
3. **DuplicationOrDeletion** - indicator of duplication or deletion for CNA.
4. **Chr** - name of a chromosome.
5. **Reference_start** - the reference position of the CNA start.
6. **Reference_end** - the reference position of the CNA end.
7. **Gene_name** - the name(s) of a gene(s).
8. **MalfunctionedByCNA** - logical indicator of whether or not the gene(s) is malfunctioned by the CNA. 
9. **mut_order** - indicator of mutation order in the simulation, it's used to detect order of mutations in the clone at each chromosome.

---

 



```{r, echo=FALSE, results='asis'}
x <- read.csv(file = "../inst/extdata/Sim_monitoring.txt",header = TRUE, sep = "\t", stringsAsFactors = FALSE)
x = x[ 1 : 20, ]
knitr::kable(x, align = "c", caption = "**Sim_monitoring.txt** file with data getting during a simulation.")
```

1. **Time** - time step. 
2. **N_clones** - total number of clone at a time step. 
3. **N_normal** - total number of normal cells at a time step. 
4. **N_primary** - total number of primary tumor cells at a time step. 
5. **N_metastatic** - total number of metastatic cells at a time step. 
6. **N_point_mutations** - total number of unique point mutations at a time step. 
7. **N_duplications** - total number of unique duplications  at a time step. 
8. **N_deletions** - total number of unique deletions at a time step. 
							
---

